### 2024-02-13 outputs/paper/paper.qmd

---
title: "Research on The Impact of Past Behaviour Normality on Regret"
subtitle: "The replication of two major experiments"
author: 
  - Ziheng Zhong
  - Yihang Cai
thanks: "Code and data are available at: https://github.com/iJustinn/Impact_of_Past_Behavior_Normality_on_Regret.git"
date: today
date-format: long
abstract: "To be written after we finished the Discussion section"
format: pdf
toc: true
toc-depth: 3
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(janitor)
library(dplyr)
library(readr)
library(knitr)
```

# Introduction

Descriptive norms and injunctive norms are key concepts in understanding social behavior, each describing different aspects of how individuals perceive and are influenced by the norms within their community or society. Descriptive norms are about observing and following what most people do in a given situation, essentially reflecting the common actions or behaviors within a group. Injunctive norms, on the other hand, are about what people believe they should or shouldn't do, based on the group's values or expectations for behavior. These norms carry a sense of approval or disapproval, guiding behavior by suggesting what is socially acceptable or frowned upon. Both types of norms are powerful in shaping individual actions by providing a roadmap of 'what is done' and 'what ought to be done,' helping maintain social order and cohesion by aligning personal conduct with group expectations.

This study delves into how people make decisions, feel about their choices, and, how they deal with regret when things don't go as planned. How much more would people regret their decisions just because they're not what you usually do? This idea stems from a concept called norm theory, which suggests that people feel extra bad about outcomes when they result from actions that are not typical for them. Two main experiments was introduced to unravel this complex bundle of decisions and emotions. The first experiment places people in a scenario where they decide whether or not to pick up a hitchhiker, testing how much one would regret the decision based on whether it's something they normally do. The second experiment changes the setting to driving, where choosing a different route leads to an accident, and looks at how this break from routine affects their feelings of regret.

By picking apart these scenarios, this paper aims to shed light on the tricky relationship between stepping out of one's comfort zone in decision-making and the regret that can follow, especially when things don't turn out well. The goal of this research is to contribute to the broader conversation on how people's minds work in these situations, improving understanding of why people feel the way they do after making certain choices. This insight is not just for psychologists but for anyone interested in the little decisions people make every day and how they shape their feelings and lives. With all been said, we predicted that someone who often picks up hitchhikers would be seen as doing something unusual (against descriptive norms) and might be judged more harshly by others (against injunctive norms). We also introduced a new aspect to the original experiment, looking at how these social norms and a person's usual actions stir up feelings, both good and bad.

The remainder of this paper is structured as follows.

@sec-data
@sec-res
@sec-dis

# Data {#sec-data}

Data used in this paper was cleaned and processed with the programming language R [@citeR]. Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `dplyr` [@citeDplyr], `readr` [@citeReadr], `knitr` [@citeknitr].

## Source

The data used in this paper was a replication of experimental data presented in the @citeOriginalPaper paper.

## Method

### Experiment #1

### Experiment #2

## Attibutes

### Experiment #1

### Experiment #2

# Results {#sec-res}

## Experiment #1

## Experiment #2

# Discussion {#sec-dis}

## First discussion point

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# References