### 2024-02-14 outputs/paper/paper.qmd
introduction modified

---
title: "Research on The Impact of Past Behaviour Normality on Regret"
subtitle: "The replication of two major experiments"
author: 
  - Ziheng Zhong
  - Yihang Cai
thanks: "Code and data are available at: https://github.com/iJustinn/Impact_of_Past_Behavior_Normality_on_Regret.git"
date: today
date-format: long
abstract: "To be written after we finished the Discussion section"
format: pdf
toc: true
toc-depth: 3
number-sections: true
bibliography: references.bib
---

```{r}
#| message: false
#| echo: false

# load packages
library(tidyverse)
library(ggplot2)
library(janitor)
library(dplyr)
library(readr)
library(knitr)
library(kableExtra)
library(reshape2)
```

```{r}
#| message: false
#| echo: false

# read in data
file_path <- "../../inputs/data/raw_data.csv"
data <- read_csv(file_path, show_col_types = FALSE)
```

# Introduction

Regret, defined as an emotional response to the memory of past decisions or actions that one wishes had been different, significantly impacts our daily lives. This paper delves into the factors that influence the intensity of regret, with a specific focus on social norms and past behaviors. Understanding these elements is essential for a clearer grasp of the decision-making process and the emotions that result from it.

We categorize social norms into two main types: descriptive norms, which concern beliefs about common behaviors within a group, and injunctive norms, which relate to beliefs about what behaviors are approved or disapproved by others. Our research involves two experiments aimed at assessing the effect of these norms on regret. The first experiment looks into the role of these norms in the context of deciding whether or not to pick up a hitchhiker, while the second addresses the consequences of selecting an unconventional driving route.

The findings from these experiments reveal that unusual or unexpected events tend to provoke stronger feelings of regret than routine behaviors. However, the influence of both injunctive and descriptive social norms on the level of regret was found to be less significant. These results highlight how deviations from usual behavior, particularly when in conflict with social norms, can intensify feelings of regret. This insight is crucial for a better understanding of the cognitive processes involved in evaluating decisions and the emotional aftermath.

The organization of this paper is methodical and clear. We start with an overview of the theoretical framework concerning social norms and their connection to regret in the introduction. The methodology and findings from our experiments are detailed in the subsequent data section. The discussion section then addresses the implications of our results, outlines the limitations of the study, and suggests directions for future research.

# Data {#sec-data}

Data used in this paper was cleaned and processed with the programming language R [@citeR]. Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `dplyr` [@citeDplyr], `readr` [@citeReadr], `knitr` [@citeknitr].

## Source

The data used in this paper was a replication of experimental data presented in the @citeOriginalPaper paper.

## Method

### Experiment #1

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-experiment1
#| tbl-cap: Result of experiment 1

# Filter the data to drop NA values
data_filtered_1 <- data %>%
  filter(!is.na(Sc1_regret) & !is.na(sc1_socnorms1) & !is.na(sc1_socnorms2) & !is.na(sc1_combinednorms))

# Calculate the total counts and counts for Mr.Jones (1) and Mr.Smith (2)
total_counts_1 <- nrow(data_filtered_1)
mr_jones_counts <- sum(data_filtered_1$Sc1_regret == 1)
mr_smith_counts <- total_counts_1 - mr_jones_counts

# Calculate the counts and percentages for Mr.Jones
mr_jones_data <- data.frame(
  Group = "Mr.Jones(exception)",
  Regret = paste(mr_jones_counts, "(", round((mr_jones_counts / total_counts_1) * 100, 1), "%)", sep = ""),
  Injunctive = paste(sum(data_filtered_1$sc1_socnorms1 == 1), "(", round((sum(data_filtered_1$sc1_socnorms1 == 1) / total_counts_1) * 100, 1), "%)", sep = ""),
  Descriptive = paste(sum(data_filtered_1$sc1_socnorms2 == 1), "(", round((sum(data_filtered_1$sc1_socnorms2 == 1) / total_counts_1) * 100, 1), "%)", sep = ""),
  `Negative Effect` = paste(sum(data_filtered_1$sc1_combinednorms == 1), "(", round((sum(data_filtered_1$sc1_combinednorms == 1) / total_counts_1) * 100, 1), "%)", sep = "")
)

# Calculate the counts and percentages for Mr.Smith
mr_smith_data <- data.frame(
  Group = "Mr.Smith(routine)",
  Regret = paste(mr_smith_counts, "(", round((mr_smith_counts / total_counts_1) * 100, 1), "%)", sep = ""),
  Injunctive = paste(sum(data_filtered_1$sc1_socnorms1 == 2), "(", round((sum(data_filtered_1$sc1_socnorms1 == 2) / total_counts_1) * 100, 1), "%)", sep = ""),
  Descriptive = paste(sum(data_filtered_1$sc1_socnorms2 == 2), "(", round((sum(data_filtered_1$sc1_socnorms2 == 2) / total_counts_1) * 100, 1), "%)", sep = ""),
  `Negative Effect` = paste(sum(data_filtered_1$sc1_combinednorms == 2), "(", round((sum(data_filtered_1$sc1_combinednorms == 2) / total_counts_1) * 100, 1), "%)", sep = "")
)

# Combine the data for Mr. Jones and Mr. Smith into one dataframe
combined_data <- rbind(mr_jones_data, mr_smith_data)

# Now use kable to create the table with the combined data
kable_output <- kable(combined_data, "latex", booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, font_size = 7)

kable_output
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-experiment1bar
#| fig-cap: Bar plot of experiment 1

smith_regret_p <- round((mr_smith_counts / total_counts_1) * 100, 1)
smith_injunctive_p <- round((sum(data_filtered_1$sc1_socnorms1 == 2) / total_counts_1) * 100, 1)
smith_descriptive_p <- round((sum(data_filtered_1$sc1_socnorms2 == 2) / total_counts_1) * 100, 1)
smith_negative_p <- round((sum(data_filtered_1$sc1_combinednorms == 2) / total_counts_1) * 100, 1)
jones_regret_p <- round((mr_jones_counts / total_counts_1) * 100, 1)
jones_injunctive_p <- round((sum(data_filtered_1$sc1_socnorms1 == 1) / total_counts_1) * 100, 1)
jones_descriptive_p <- round((sum(data_filtered_1$sc1_socnorms2 == 1) / total_counts_1) * 100, 1)
jones_negative_p <- round((sum(data_filtered_1$sc1_combinednorms == 1) / total_counts_1) * 100, 1)

# Create a data frame with the provided data
df <- data.frame(
  Category = c("Regret", "Injunctive Social Norm", "Descriptive Social Norm", "Negative Affect"),
  Smith_Routine = c(smith_regret_p, smith_injunctive_p, smith_descriptive_p, smith_negative_p),
  Jones_Exception = c(jones_regret_p, jones_injunctive_p, jones_descriptive_p, jones_negative_p)
)

# Convert the data from wide to long format
data_long <- melt(df, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")

# Reorder the factor levels of Category so that 'Regret' comes first
data_long$Category <- factor(data_long$Category, levels = c("Regret", "Injunctive Social Norm", "Descriptive Social Norm", "Negative Affect"), 
                             labels = c("Regret", "Social Norm\nInjunctive", 
                                        "Social Norm\nDescriptive", "Negative\nAffect"))

# Create the bar plot with the categories reordered
ggplot(data_long, aes(x = Category, y = Percentage, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.7) +
  scale_fill_manual(values = c("Smith_Routine" = "grey", "Jones_Exception" = "#505050"), 
                    labels = c("Mr.Smith(Routine)", "Mr.Jones(Exception)")) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20)) +
  theme_minimal() +
  labs(x = "", y = "Percentage", fill = "Condition") +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, colour = "black",size = 9), 
    legend.position = "top", legend.title = element_blank())
```

### Experiment #2

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-experiment2
#| tbl-cap: Result of experiment 2

# Filter the data to drop NA values
data_filtered_2 <- data %>%
  filter(!is.na(Sc2_regret) & !is.na(Sc2_random_1) & !is.na(Sc2_random_2) & !is.na(Sc2_lucky))

# Calculate the total counts and counts for Mr.Adams (1) and Mr.White (2)
total_counts_2 <- nrow(data_filtered_2)
mr_adams_counts <- sum(data_filtered_2$Sc2_regret == 1)
mr_white_counts <- total_counts_2 - mr_adams_counts

# Calculate the counts and percentages for Mr.Adams
mr_adams_data <- data.frame(
  Group = "Mr.Adams(routine)",
  Regret = paste(mr_adams_counts, "(", round((mr_adams_counts / total_counts_2) * 100, 1), "%)", sep = ""),
  Luck = paste(sum(data_filtered_2$Sc2_lucky == 1), "(", round((sum(data_filtered_2$Sc2_lucky == 1) / total_counts_2) * 100, 1), "%)", sep = "")
)

# Calculate the counts and percentages for Mr. White
mr_white_data <- data.frame(
  Group = "Mr.White(exception)",
  Regret = paste(mr_white_counts, "(", round((mr_white_counts / total_counts_2) * 100, 1), "%)", sep = ""),
  Luck = paste(sum(data_filtered_2$Sc2_lucky == 2), "(", round((sum(data_filtered_2$Sc2_lucky == 2) / total_counts_2) * 100, 1), "%)", sep = "")
)

# Combine the data for Mr. Jones and Mr. Smith into one dataframe
combined_data <- rbind(mr_adams_data, mr_white_data)

# Now use kable to create the table with the combined data
kable_output <- kable(combined_data, "latex", booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, font_size = 7)

kable_output
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-experiment2bar
#| fig-cap: Bar plot of experiment 2

adams_regret_p <- round((mr_adams_counts / total_counts_2) * 100, 1)
adams_luck_p <- round((sum(data_filtered_2$Sc2_lucky == 1) / total_counts_2) * 100, 1)
white_regret_p <- round((sum(data_filtered_2$Sc2_regret == 2) / total_counts_2) * 100, 1)
white_luck_p <- round((sum(data_filtered_2$Sc2_lucky == 2) / total_counts_2) * 100, 1)

# Create a data frame with the provided data
data <- data.frame(
  Category = c("Regret", "Luck"),
  Adams_Routine = c(adams_regret_p, adams_luck_p),
  White_Exception = c(white_regret_p, white_luck_p)
)

# Convert the data from wide to long format
data_long <- melt(data, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")

# Reorder the factor levels of Category so that 'Regret' comes first
data_long$Category <- factor(data_long$Category, levels = c("Regret", "Luck"))

# Create the bar plot with the categories reordered
ggplot(data_long, aes(x = Category, y = Percentage, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.7) +
  scale_fill_manual(values = c("Adams_Routine" = "grey", "White_Exception" = "#505050"), 
                    labels = c("Mr.Adams(Routine)", "Mr.White(Exception)")) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20)) + # label the y-axis every 20%
  theme_minimal() +
  labs(x = "", y = "Percentage", fill = "Condition") +
  theme(axis.text.x = element_text(hjust = 1, vjust = 1, colour = "black",size = 9), 
    legend.position = "top", legend.title = element_blank())
```

## Attibutes

### Experiment #1

### Experiment #2

# Results {#sec-res}

## Experiment #1

## Experiment #2

# Discussion {#sec-dis}

## First discussion point

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# References